Public Support for Democracy, Regime Change, and the Importance of Incorporating Measurement Uncertainty

### 2020-11-16 paper/dcpo_demsupport.Rmd
make Hu corresponding author; revise some citations

---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: svm-latex-ms2.tex
title: "Public Support for Democracy, Regime Change, and the Importance of Incorporating Measurement Uncertainty"
thanks: "The authors contributed equally to this work.  Corresponding author: [yuehong-tai@uiowa.edu](mailto:yuehong-tai@uiowa.edu).  Current version: `r format(Sys.time(), '%B %d, %Y')`."
author:
- name: Yuehong 'Cassandra' Tai
  affiliation: University of Iowa
- name: Yue Hu
  affiliation: Tsinghua University
- name: Frederick Solt
  affiliation: University of Iowa
abstract: "Do democratic regimes depend on public support to avoid backsliding? Does public support, in turn, respond thermostatically to changes in regime? Two recent prominent studies (Claassen 2020a, 2020b) have reinvigorated the classic hypothesis on the positive relationship between public support for democracy and regime survival--and challenged its reciprocal counterpart--by using a Bayesian latent variable approach to measure mass democratic support from cross-national survey data. But such approaches come with concomitant measurement uncertainty, and neither study incorporated this error into its analyses. Here we correctly take this uncertainty in account and show that there is no support for the conclusion of either study. We then work to minimize the measurement uncertainty in public support by bringing additional survey data and a superior model of public opinion. Even with these improvements, however, our analyses fail to yield evidence in support of either hypothesis."
keywords: ""
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
spacing: double
bibliography: \dummy{`r file.path(getwd(), list.files(getwd(), ".bib$", recursive = TRUE))`}
# csl: https://raw.githubusercontent.com/citation-style-language/styles/master/american-political-science-association.csl
biblio-style: apsr
citecolor: black
linkcolor: black
endnote: no
header-includes:
      - \usepackage{array}
      - \usepackage{caption}
      - \usepackage{graphicx}
      - \usepackage{siunitx}
      - \usepackage{colortbl}
      - \usepackage{multirow}
      - \usepackage{hhline}
      - \usepackage{calc}
      - \usepackage{tabularx}
      - \usepackage{threeparttable}
      - \usepackage{wrapfig}
nocite: |
  @Solt2020a
---
# Introduction

One of the classic but disputable questions in comparative politics is the relationship between mass democratic support and countries' level of democracy.
A long-standing and widely accepted theory is that the level of mass democratic support determines the stability of a democratic system [e.g., @lipset_social_1959; @norris_democratic_2011], but findings from empirical analyses still remain inconclusive [@inglehart_modernization_2005; @fails_changing_2010; @qi_how_2011; for a review in this literature, see @Claassen2020a].

@@

In our reexamination, instead of employing the point estimates, we followed the recommendation and the approach of incorporating uncertainty in using latent variables [e.g., @Schnakenberg2014; @Crabtree2015] and other measures, such as the Standardized World Income Inequality Database [@Solt2020].
Specifically, in our replication of Claassen's [-@Claassen2020a] analysis models, we incorporated uncertainty by using random draws from the posterior distributions. 
We generated 1,000 duplicate versions of the analysis dataset, assigned to each a different random draw from the posterior distributions of public democratic support, performed the analyses repeatedly on each of these 1,000 versions of the dataset, and then combined the results following the rules set out in @Rubin1987.

@@

As previous operation, we tested if this alteration is caused by the data scarcity or estimating inefficiency. 
When more data and better method are applied, the confidence intervals do shrink yet remaining statistically insignificant. 
We observed the similar changes after adjusting the effect of domestic corruption [@Claassen2020b: 41].

@@

In addition to the potentially methodological cause (the measurement errors), there are some potentially substantive causes for the disappearing relationship.
First, democratic backsliding might be mainly caused by elites.
Elites are thought to play a critical role in preserving democratic stability by respecting the limits on their behaviors [@weingast_political_1997].
After the Cold War, many democracies backslid and were imperiled worldwide because elected leaders lack commitment to democratic norms when the public even did not notice that [@bermeo_democratic_2016,@levitsky_how_2018, @waldner_unwelcome_2018].

@@

# Conclusion
In this note, we reexamined the findings from Claassens'[-@Claassen2019; -@Claassen2020b] paper that public support helps the survival of democracy and democratic development has a thermostatic effect on public support for democracy.
We demonstrated that the relationships between public support and democratic development depends on the incorporation of uncertainty in the measurement of latent variables.
Even using more data and a sophisticated DCPO model, the measurement uncertainty made the relationships irreplicable irrespective of whether public support is an independent variable or dependent variable.

>>>>

---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: svm-latex-ms2.tex
title: "Public Support for Democracy, Regime Change, and the Importance of Incorporating Measurement Uncertainty"
thanks: "The authors contributed equally to this work.  Corresponding author: [yuehu@mail.tsinghua.edu.cn](mailto:yuehu@mail.tsinghua.edu.cn).  Current version: `r format(Sys.time(), '%B %d, %Y')`."
author:
- name: Yuehong 'Cassandra' Tai
  affiliation: University of Iowa
- name: Yue Hu
  affiliation: Tsinghua University
- name: Frederick Solt
  affiliation: University of Iowa
abstract: "Do democratic regimes depend on public support to avoid backsliding? Does public support, in turn, respond thermostatically to changes in regime? Two prominent recent studies (Claassen 2020_a_, Claassen 2020_b_) have reinvigorated the classic hypothesis on the positive relationship between public support for democracy and regime survival--and challenged its reciprocal counterpart--by using a Bayesian latent variable approach to measure mass democratic support from cross-national survey data. But such approaches come with concomitant measurement uncertainty, and neither study incorporated this uncertainty into its analyses. Here we correctly take this uncertainty in account and show that there is no support for the conclusion of either study. We then work to minimize the measurement uncertainty in public support by bringing additional survey data and a superior model of public opinion. Even with these improvements, however, our analyses fail to yield evidence in support of either hypothesis. [142/150 words]"
keywords: ""
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
spacing: double
bibliography: \dummy{`r file.path(getwd(), list.files(getwd(), ".bib$", recursive = TRUE))`}
# csl: https://raw.githubusercontent.com/citation-style-language/styles/master/american-political-science-association.csl
biblio-style: apsr
citecolor: black
linkcolor: black
endnote: no
header-includes:
      - \usepackage{array}
      - \usepackage{caption}
      - \usepackage{graphicx}
      - \usepackage{siunitx}
      - \usepackage{colortbl}
      - \usepackage{multirow}
      - \usepackage{hhline}
      - \usepackage{calc}
      - \usepackage{tabularx}
      - \usepackage{threeparttable}
      - \usepackage{wrapfig}
---
# Introduction

One of the classic but disputable questions in comparative politics is the relationship between mass democratic support and countries' level of democracy.
A long-standing and widely accepted theory is that the level of mass democratic support determines the stability of a democratic system [e.g., @lipset_social_1959; @norris_democratic_2011], but findings from empirical analyses still remain inconclusive [@inglehart_modernization_2005; @fails_changing_2010; @qi_how_2011; for a review of this literature, see @Claassen2020a].

@@

In our reexamination, instead of employing the point estimates, we followed the recommendation and the approach of incorporating uncertainty in using latent variables [e.g., @Schnakenberg2014; @Crabtree2015].
Specifically, in our replication of Claassen's [-@Claassen2020a] analysis models, we incorporated uncertainty by using random draws from the posterior distributions. 
We generated 1,000 duplicate versions of the analysis dataset, assigned to each a different random draw from the posterior distributions of public democratic support, performed the analyses repeatedly on each of these 1,000 versions of the dataset, and then combined the results following the rules set out in @Rubin1987.

@@

As previous operation, we tested if this alteration is caused by the data scarcity or estimating inefficiency. 
When more data and better method are applied, the confidence intervals do shrink yet remaining statistically insignificant. 
We observed the similar changes after adjusting the effect of domestic corruption [@Claassen2020b, 41].

@@

In addition to the potentially methodological cause (the measurement errors), there are some potentially substantive causes for the disappearing relationship.
First, democratic backsliding might be mainly caused by elites.
Elites are thought to play a critical role in preserving democratic stability by respecting the limits on their behaviors [@weingast_political_1997].
After the Cold War, many democracies backslid and were imperiled worldwide because elected leaders lack commitment to democratic norms when the public even did not notice that [@bermeo_democratic_2016; @levitsky_how_2018; @waldner_unwelcome_2018].

@@

# Conclusion
In this note, we reexamined the findings from Claassens' [-@Claassen2020a; -@Claassen2020b] paper that public support helps the survival of democracy and democratic development has a thermostatic effect on public support for democracy.
We demonstrated that the relationships between public support and democratic development depends on the incorporation of uncertainty in the measurement of latent variables.
Even using more data and a sophisticated DCPO model, the measurement uncertainty made the relationships irreplicable irrespective of whether public support is an independent variable or dependent variable.
